### 2018-11-08 paper.tex
Initial commit

% MATH 5251 Final paper.

\documentclass[12pt]{article}

\usepackage{graphicx}
\usepackage{amssymb, amsmath, amsthm, amsfonts}
\usepackage{tikz}

\textwidth = 7 in
\textheight = 9.5 in
\oddsidemargin = -0.3 in
\evensidemargin = -0.3 in
\topmargin = -0.4 in
\headheight = 0.0 in
\headsep = 0.0 in
\parskip = 0.2in
\parindent = 0.0in

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}{Proposition}
\newtheorem{defn}{Definition}[section]

\title{A Review of Data Compression Methods}
\author{Anonymous}

\begin{document}
\maketitle

In the age of ever-expanding quantities of digital data, the ability
to store this data efficiently is more important than ever. Throughout
the history of the field of computer science, various algorithms of
varying complexities and efficacies have been created in order to
reduce the amount of space required to store arbitrary strings of
binary data. This wide variety of methods are optimized for different
purposes: some algorithms attempt to reduce the amount of space
required to store raster images; some try to optimize audio storage;
some can optimize genetic data to an extremely high degree; and still
more attempt to work well for general-purpose use. Within the scope of
this paper, I will attempt to give an overview of a handful of popular
compression schemes, including general purpose as well domain-specific
methods, as well as a theoretical overview of the limits and potential
benefits of compression in general. This paper will only cover
lossless data encoding: that is, alogrithms which can reproduce a
perfect copy of the input data from its output. Lossy compression
schemes exist which can achieve incredible compression ratios, but
their methods typically include psychological ``tricks'' which allows
the algorithm to discard data which is less noticable to human
senses. As such, these methods are outside the scope of this paper.

\section{Theoretical Limits to Data Compression}

Data compression has many benefits as quickly reviewed above; however,
many limitations also exist. Note that for simplicity, we will limit
our discussion of compression schemes to binary schemes, as these are
the most wide spread in applications in the real world (i.e.,
compression of computer files.)

\begin{thm}
  For any given data compression scheme, there exist some input
  whose output is not shorter than its input. That is to say, there is
  no compression scheme which can compress every input bit-string.
\end{thm}

\begin{proof}
  Informally, we can describe a ``perfect compression scheme'' as some
  mapping which uniquely maps a given input bit-string to an output
  bit-string of strictly lesser length. More formally, a perfect
  compression scheme can be modeled as a function which maps input
  bit-strings to output bitstrings which are shorter. For any
  $n \in \mathbb{N}$, let $S_n$ be the set of bit-strings of length
  $n$, and let $T_n = \bigcup_{k = 1}^{n - 1} S_k$; that is, $T_n$ is
  the ``target'' set of bit-strings which are of lengths are strictly
  less than $n$. Suppose that there exists some ``perfect''
  compression scheme $C:S_n \to T_n$. The set $S_n$ has $2^n$ elements
  and $T_n$ has $\sum_{k=1}^{n-1}2^k = 2^n - 2$ elements. Since
  $2^n > 2^n - 2$, by the pigeonhole principle there must be at least
  one element in $T_n$ such that two distinct elements in $S_n$ map to
  this element through $C$. However, $C$ is defined to be a function,
  so this is a contradiction, and thus such a function $C$ can not
  exist.
\end{proof}

This result is not to say that compression is impossible; on the
contrary! It is very possible for a compression scheme to reduce
\emph{on average} the length of its inputs.

\section{Huffman Codes}

\begin{thm}
  This is a theorem.
\end{thm}

\begin{proof}
This is a proof.
\end{proof}

The equation $y=x$ is an in-line equation.

The equation
\[
y=x
\]
is a display style equation.

\section{Lempel-Ziv Based Methods}

If there's anything you don't know how to do with \LaTeX{}, try
googling it. If you still can't figure it out, email me.

\end{document}